\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-27 paper/colm_submission/colm2024_conference.tex
COLM draft

\documentclass[10pt]{article} % For LaTeX2e
\usepackage[preprint]{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{url}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{pdflscape} 
\usepackage{multirow} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{bbm}
\usepackage{fontawesome}
\usepackage{fancyvrb}

@@

% Custom colors
\usepackage{color}

@@

% For model colours
\definecolor{deepblue}{rgb}{0.29411765 0.45882353 0.61960784}
\definecolor{deepred}{rgb}{0.74509804 0.21176471 0.23921569}
\definecolor{deepgreen}{rgb}{0,0.5,0}
\definecolor{deeppurple}{rgb}{0.52941176 0.32941176 0.56470588}

% For code listings
\definecolor{codegreen}{rgb}{0,0.4,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

@@

\newcommand{\package}[1]{\textsl{#1}\xspace}
\newcommand{\eg}{{e.\,g.}\xspace}
\newcommand{\ie}{{i.\,e.}\xspace}
\newcommand{\hubble}{\emph{Hubble}\xspace}
\newcommand{\eqrefb}[1]{(\ref{#1})}

\newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
% Changes
\newcommand{\changes}[1]{\textcolor{blue}{#1}}

@@

\lstdefinestyle{mystyle}{ 
  backgroundcolor=\color{backcolour}, 
  commentstyle=\color{deepgreen},
  keywordstyle=\color{deepred}, 
  numberstyle=\tiny\color{codegray}, 
  stringstyle=\color{deepgreen},
  basicstyle=\ttfamily\footnotesize\linespread{1.1},
  breakatwhitespace=false, 
  breaklines=true, captionpos=b, 
  keepspaces=true, numbers=left, 
  numbersep=8pt, showspaces=false, 
  showstringspaces=false, showtabs=false, 
  frame=single, 
  framerule=0.2pt, 
  rulecolor=\color{codegray}, 
  tabsize=2, 
  aboveskip=1.5ex, 
  belowskip=1.5ex,
  xleftmargin=15pt, 
  xrightmargin=15pt, 
  extendedchars=true, 
  columns=flexible, 
  linewidth=\textwidth 
  }

@@

% Define Verbatim environment with custom style
\DefineVerbatimEnvironment{jsoncode}{Verbatim}{
  commandchars=\\\{\},
  rulecolor=\color{codegray},
  fillcolor=\color{codegray},
  labelposition=topline,
  fontsize=\small,
  baselinestretch=1.1,
  formatcom=\color{codegreen},
  xleftmargin=15pt,
  xrightmargin=15pt,
  tabsize=2
}

@@

\author{\name Siddharth Mishra-Sharma \email \href{mailto:smsharma@mit.edu}{smsharma@mit.edu} \\
      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
      Department of Physics, Harvard University, Cambridge, MA 02138, USA
      \AND
      \name Yiding Song \email \href{mailto:ydsong@mit.edu}{ydsong@mit.edu} \\
      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
      Department of Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
      \AND
      \name Jesse Thaler \email \href{mailto:jthaler@mit.edu}{jthaler@mit.edu} \\
      \addr The NSF AI Institute for Artificial Intelligence and Fundamental Interactions\\
      Center for Theoretical Physics, Massachusetts Institute of Technology, Cambridge, MA 02139, USA \\
}

@@

\def\month{MM}  % Insert correct month for camera-ready version
\def\year{YYYY} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

\begin{document}

@@

\thispagestyle{firstpage}

\begin{abstract}
We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by surveys and telescopes with natural language  using a neural network model. The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts, optionally summarized via guided generation using large language models (LLMs), and corresponding downstream observations. Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and text through tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries) and description retrieval (i.e., querying for astrophysical object classes and science use cases most relevant to a given observation). Our study demonstrates the potential for using generalist \changes{foundation models} rather than task-specific models for \changes{interacting with} astronomical data by leveraging text as an interface. \githubmaster
\end{abstract}

\tableofcontents

@@

Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
%
ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
%
Until recently, most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
%
The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
%
These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
%
Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
%

In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations. 
%
Concretely, we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm. \changes{A high-level overview of the method is shown in Fig.~\ref{fig:overview}}.
%

The rest of this paper is organized as follows.
%
In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
%
In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
%
In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on retrieval tasks.
%
We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.

\begin{figure*}[!t]
\centering
\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
\caption{\changes{Overview of the PAPERCLIP method. (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts. The proposal abstracts are optionally summarized using guided large language model generation. (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query). The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
\label{fig:overview}
\end{figure*}

\section{\changes{Related Work}}

% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
%
The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}. For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
%

Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
%

\citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.

@@

% While the interaction of astrophysical data with large language models is fairly nascent, \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is a recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.

In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observational proposals to inform text captions.
%

@@

Table \ref{tab:dataset} shows examples of images and their corresponding (clipped) proposal abstracts.
%
It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models \citep{deng2009imagenet}.
%
This further motivates the need for fine tuning on domain-specific data.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{7.5cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column). The observation cycle and corresponding year, as well as proposal ID, are shown in the second and third columns, respectively. The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
  \label{tab:dataset}
\end{table}

\begin{table}[h!]
  \renewcommand{\arraystretch}{2}
    \centering
    \begin{tabular}{m{1.8cm} m{6.6cm} m{6.6cm}}
        \toprule
        \bfseries Prop. ID & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
        \cmidrule(r){2-3}
        & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
        \midrule
        \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
        \midrule
        \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
        \midrule
        \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
        \midrule
        \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
        \bottomrule
    \end{tabular}
    \caption{\changes{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset},} the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately.}
    \label{tab:datasetsumm}
\end{table}

@@

Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} API.
%
Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
%
We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic purposes.
%
A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations and survey-style campaigns.
%
Images are centered and resized to a resolution-per-side of 512 pixels.
%
Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the dataset; models trained on datasets with color images included were observed to show worse performance on evaluation metrics.
%
If no appropriate images corresponding to an abstract are found, it is excluded from the dataset.

In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
%
3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation. The held out images correspond to 438 unique abstracts.  % Double check numbers in final version.

We note that some fraction of the image-caption pairs in the constructed dataset will primarily concern instrumentation and/or calibration rather than scientific content.
%
We choose to not filter out these pairs from our dataset, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.

% Mention that the signal is noisy!

@@

The method from \cite{willard2023efficient} is used to produce an LLM-generated summary of the abstract conforming to a particular schema, specified in JSON format.
%
The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.

The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the generation process by masking out tokens that would violate the intended format.
%
By framing the problem in terms of transitions between a set of finite states (a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
%
See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here.

While the schema-guided generation ensures the \emph{format} of the output, the prompt and choice of LLM will dictate the \emph{content} of the generated summaries.
%
We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}~\citep{jiang2024mixtral} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
%
Further details on the summarization procedure, including the prompts and schema used, are provided in App.~\ref{app:summarization}.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{colm2024_conference}

\usepackage{microtype}
% \usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}

\usepackage{graphicx}
\usepackage{array}
\usepackage{xspace}
\usepackage{pdflscape} 
\usepackage{multirow} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{bbm}
\usepackage{fontawesome}
\usepackage{fancyvrb}

@@

\definecolor{darkblue}{rgb}{0, 0, 0.5}
% \hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue, urlcolor=darkblue}

@@

% For model colours and code
\definecolor{deepblue}{rgb}{0.29411765 0.45882353 0.61960784}
\definecolor{deepred}{rgb}{0.74509804 0.21176471 0.23921569}
\definecolor{deepgreen}{rgb}{0,0.5,0}
\definecolor{deeppurple}{rgb}{0.52941176 0.32941176 0.56470588}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

@@

\newcommand{\package}[1]{\textsl{#1}\xspace}
\newcommand{\hubble}{\emph{Hubble}\xspace}
\newcommand{\eqrefb}[1]{(\ref{#1})}

\newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
\newcommand{\changes}[1]{\textcolor{blue}{#1}}

@@

% Listings style
\lstdefinestyle{mystyle}{ 
  backgroundcolor=\color{backcolour}, 
  commentstyle=\color{deepgreen},
  keywordstyle=\color{deepred}, 
  numberstyle=\tiny\color{codegray}, 
  stringstyle=\color{deepgreen},
  basicstyle=\ttfamily\footnotesize\linespread{1.1},
  breakatwhitespace=false, 
  breaklines=true, captionpos=b, 
  keepspaces=true, numbers=left, 
  numbersep=8pt, showspaces=false, 
  showstringspaces=false, showtabs=false, 
  frame=single, 
  framerule=0.2pt, 
  rulecolor=\color{codegray}, 
  tabsize=2, 
  aboveskip=1.5ex, 
  belowskip=1.5ex,
  xleftmargin=15pt, 
  xrightmargin=15pt, 
  extendedchars=true, 
  columns=flexible, 
  linewidth=\textwidth 
  }

@@

% Define Verbatim environment with custom style
\DefineVerbatimEnvironment{jsoncode}{Verbatim}{
  commandchars=\\\{\},
  rulecolor=\color{codegray},
  fillcolor=\color{codegray},
  labelposition=topline,
  fontsize=\small,
  baselinestretch=1.1,
  formatcom=\color{deepgreen},
  xleftmargin=15pt,
  xrightmargin=15pt,
  tabsize=2
}

@@

% Authors must not appear in the submitted version. They should be hidden
% as long as the \colmfinalcopy macro remains commented out below.
% Non-anonymous submissions will be rejected without review.

\author{Antiquus S.~Hippocampus, Natalia Cerebro \& Amelie P. Amygdale \thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.  Funding acknowledgements go at the end of the paper.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213, USA \\
\texttt{\{hippo,brain,jen\}@cs.cranberry-lemon.edu} \\
\And
Ji Q. Ren \& Yevgeny LeNet \\
Department of Computational Neuroscience \\
University of the Witwatersrand \\
Joburg, South Africa \\
\texttt{\{robot,net\}@wits.ac.za} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email}
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

@@

% \colmfinalcopy % Uncomment for camera-ready version, but NOT for submission.
\begin{document}

@@

\begin{abstract}
   We present PAPERCLIP (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training), a method which associates astronomical observations imaged by telescopes with natural language using a neural network model. The model is fine-tuned from a pre-trained Contrastive Language–Image Pre-training (CLIP) model using successful observing proposal abstracts and corresponding downstream observations, with the abstracts optionally summarized via guided generation using large language models (LLMs). Using observations from the \hubble Space Telescope (HST) as an example, we show that the fine-tuned model embodies a meaningful joint representation between observations and natural language through quantitative evaluation as well as tests targeting image retrieval (i.e., finding the most relevant observations using natural language queries).
  and description retrieval (i.e., querying for astrophysical object classes and use cases most relevant to a given observation). 
  Our study demonstrates the potential for using generalist foundation models rather than task-specific models for interacting with astronomical data by leveraging text as an interface.
\end{abstract}

@@

Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
%
ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
%
Most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
%
The {foundation model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
%
These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream.
%
Foundation models can often benefit from additional training ({fine-tuning}) using a relatively small amount of domain-specific data in order to increase their usefulness when applied to specialized domains.

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards,mccabe2023multiple,Birk:2024knn,vig2024finetuning,heinrich2024masked}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse ways of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
% This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint physically-informative embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
% %

In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training),
% \footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), 
a method that connects, for the first time, astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals written by astronomers and images corresponding to downstream observations imaged by telescopes. 
%
Concretely, we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm. A high-level overview of the method is shown in Fig.~\ref{fig:overview}.
%

\begin{figure*}[!t]
   \centering
   \includegraphics[width=0.97\textwidth]{plots/figure.pdf}
   \caption{Overview of the PAPERCLIP method. (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts. The proposal abstracts are optionally summarized using guided large language model generation. (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval i.e., finding the observations most relevant to a given text query. The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}.}
   \label{fig:overview}
   \end{figure*}

  %  \section{Related Work}
  %  \label{sec:related}

   \paragraph*{Related Work}

   The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}. For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
   %
   Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
   %
   \citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.
  %  %
  In contrast with previous work, our application is the first to associate astronomical observation with the text modality in a generalist framework, showcasing the potential of language models in specialized scientific domains like astronomy.
  %

@@

  %  In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observing proposal abstracts to inform text captions.
  %  %

  The rest of this paper is organized as follows.
%
% We review related work in Sec.~\ref{sec:related}.
% %
In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
%
In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
%
In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on retrieval tasks.
%
We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.

@@

   Table \ref{tab:dataset} shows examples of images and their corresponding (clipped) proposal abstracts.
   %
   It can be seen that the images in this dataset exhibit specific characteristics as well as artifacts particular to HST data-taking and processing which distinguishes them from the distribution of natural images typically used for large-scale pre-training of foundation models.
   %
   This further motivates fine-tuning on domain-specific data.

@@

   \begin{table}[h!]
     \centering
     \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{5.5cm}}
         \toprule
         \centering \bfseries \hubble image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
         \midrule
         \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
         \midrule
        %  \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
        %  \midrule
         \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
        %  \midrule
        %  \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
         \bottomrule
     \end{tabular}
     \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column). The observation cycle and corresponding year, as well as proposal ID, are shown in the second and third columns, respectively. The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
     \label{tab:dataset}
   \end{table}

   \begin{table}[h!]
     \renewcommand{\arraystretch}{2}
       \centering
       \begin{tabular}{m{1.8cm} m{3.6cm} m{7.2cm}}
           \toprule
           \bfseries Prop. ID & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
           \cmidrule(r){2-3}
           & \centering\arraybackslash \bfseries \small{Objects and phenomena} & \centering\arraybackslash \bfseries \small{Science use cases} \tabularnewline
           \midrule
           \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
           \midrule
          %  \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
          %  \midrule
           \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
          %  \midrule
          %  \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
           \bottomrule
       \end{tabular}
       \caption{For the \hubble proposal abstracts shown in Tab.~\ref{tab:dataset}, the LLM (\textsc{Mixtral-8x7B})-extracted summaries showing objects and phenomena (middle column) as well as potential downstream science use cases (last column) separately. The proposal IDs (left column) contain hyperlinks to the MAST page corresponding to the proposal.}
       \label{tab:datasetsumm}
   \end{table}

@@

   Observations corresponding to individual proposal IDs are queried through the Mikulski Archive for Space Telescopes (MAST)\footnote{\url{https://mast.stsci.edu/}} via the \package{Astroquery} \citep{2019AJ....157...98G} API.
   %
   Products of type \texttt{PREVIEW} are filtered in, corresponding to preview postcard images.
   %
   We note that these are not science-grade observations, but rather lower-resolution images useful for diagnostic or preview purposes.
   %
   A maximum of 20 images are downloaded per proposal ID, selected at random, in order to avoid biasing the model towards proposals with a larger number of observations and survey-style campaigns.
   %
   Images are centered and resized to a resolution-per-side of 512 pixels.
   %
   Color previews (i.e., observations taken with multiple wavelength filters assigned to individual RGB channels) are manually excluded via a filename filter in order to maintain consistency across the samples.
   % models trained on datasets with color images included were observed to show worse performance on evaluation metrics.
   %
   If no appropriate images corresponding to an abstract are found, it is excluded from the dataset.

   In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
   %
   3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation. The held out images correspond to 429 unique abstracts.  % Double check numbers in final version.

   We note that some fraction of the image-caption pairs in the constructed dataset will primarily concern instrumentation and/or calibration rather than scientific content.
   %
   We choose to not filter out these pairs, in order to have a larger sample of HST observations that the model can leverage to adapt to the distinctive characteristics of \hubble images.

@@

   The method from \cite{willard2023efficient} is used to produce an LLM-generated summary of the abstract conforming to a particular schema, specified in JSON format.
   %
   The schema is designed to represent a list of the objects (e.g., `Type Ia supernova') and phenomena (e.g., `gravitational lensing'), as well as potential downstream science uses cases (e.g., `set constraints on supernova explosion models') that could correspond to the eventual imaged observation given the abstract text, with a minimum of 1 and a maximum of 5 elements per list.
   %
   The procedure guides the generation of LLM outputs while ensuring that the schema is respected at every step in the generation process by masking out tokens that would violate the intended format.
   %
   By framing the problem in terms of transitions between a set of finite states (i.e., a finite-state machine), \cite{willard2023efficient} showed that guided generation can be performed with negligible overhead compared to unconstrained generation.
   %
   See App.~\ref{app:guided-generation} for a more detailed description of the guidance generation method used here, including an overview of technical details.
   %
   While the schema-guided generation ensures the \emph{format} of the output, the prompt and choice of LLM will dictate the \emph{content} of the generated summaries.
   %
   We use the open-weights, instruction-tuned model \textsc{Mixtral-8x7B-Instruct}~\citep{jiang2024mixtral} to generate the summaries, with guided generation performed using the \package{Outlines}\footnote{\url{https://github.com/outlines-dev/outlines}} package.
   %
   Further details on the summarization procedure, including the prompts and schema used, are provided in App.~\ref{app:summarization}.
